---
title: "Money and Politics"
author: "Christine, Nybria, Yuri"
date: "March 21, 2018"
output:
  html_document:
    code_folding: hide
--- 


## Loading the data
```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
wordcountaddin::text_stats("mp2.Rmd")
```

## Initial clean up of the data 
```{r, message=FALSE, warning=FALSE, echo=TRUE}
contributions1 <- merge(candidates, contributions, by = "cand_id") # merged the candidate dataframe and contribution dataframe by the cand_id to see the name of the candidate

house_elections1 <- rename(house_elections, cand_id = fec_id) # renamed the row 

cont_winner <- contributions1 %>%
  inner_join(house_elections1, by = "cand_id") 


new_cont <- contributions1 %>%
  select(cand_id, cand_name, cand_party_affiliation,  cand_election_yr, cand_office_state , name, transaction_amt) # there are a lot of contributions to the same representative. 


contributions1 %>% group_by(cand_party_affiliation) %>% summarize(num_type = n()) %>% arrange(desc(num_type)) %>% head(4)

cont_winner %>%
  select(cand_id, cand_name, cand_party_affiliation, transaction_amt, ge_winner) %>%
  filter(transaction_amt > 0 ) %>% #remove negative transactions
  group_by(cand_name)

cont_winner %>%
  group_by(cand_party_affiliation) %>%
  summarise( nrow = n()) %>%
  arrange(desc(nrow)) %>%
  head (5) # finding the top 5 parties with the highest number of memebers in the election


# This function will be used to filter out transaction amounts and wins for each polical party 
party <- function(name_arg) {
  cont_winner %>%
    filter(cand_party_affiliation == name_arg) %>%
    filter(ge_winner == "W") %>%
    select(transaction_amt, ge_winner) %>%
    group_by(transaction_amt, ge_winner) %>%
    summarise(trans = sum(transaction_amt), win = n()) %>%
    arrange(desc(transaction_amt))
}

# This function will be used to filter out transaction amounts and losses for each polical party 
party2 <- function(name_arg) {
  cont_winner %>%
    filter(cand_party_affiliation == name_arg) %>%
    filter(ge_winner == "N") %>%
    group_by(transaction_amt, ge_winner) %>%
    summarise(trans = sum(transaction_amt), win = n()) %>%
    arrange(desc(transaction_amt))
}  

# analyzing each of of the 5 parties and their tranaction amounts 
affiliation <- c("DEM", "REP", "DFL", "UNK", "OTH")
 lapply(affiliation, FUN = party2)
  

cont_winner2 <- cont_winner %>%
  filter(cand_party_affiliation %in% c("REP", "DEM", "DFL", "UNK", "OTH"))

```

What we wanted to find was weather or not there is a relationship between money and a candidate's chances of winning. Our hypothesis was that the more money a candidate received the more likely they were to win in a election. Our first step with data wrangling creating a new data frame, `Contributions 1`, which merged candidate id with candidate name. Since we wanted to observe total transaction for each candidate, we took the next step to inner join our new data frame,  `Contributions 1` with one of the original data frames `house_elections`.

Our first step to analyzing our hypothesis, was developing a function that filtered transaction amounts (`transaction_amt`) and wins and losses (`ge_winner`), however, we concluded that finding transaction amounts for each name was too exhaustive, so we changed our focus from candidate name and transaction amount  to party affiliation and transaction amount. 

There are many factors that come into play when running for federal positions. Is there a correlation between the money and politics?  One [article] (https://www.nytimes.com/roomfordebate/2016/02/25/does-money-really-matter-in-politics/the-power-of-politcal-money-is-overrated) voices that money does not guarnteee a win in a general election. The Federal Election Commision data would give us the proper information to test our hypothesis. Datasets that we used are `house_elections`, `candidates`, and `contributions`. In order to get a workable dataset that included both the name of the candidate and the contributions that they received we joined the `contributions` and `candidate` datasets. To further analyze the relationship between money and a candidate’s election result, we used `house_elections` dataset to see if the candidate won or lost in the 2011-2012 election. By joining the house election data to our main dataset, this would give us the variables that we wanted to look at: transaction amount, party affiliation, and the election results. 

```{r, message=FALSE, warning=FALSE, echo=TRUE}
#This will change the ge_winner factors around so that win will appear as the first bar with lose following. 
cont_winner2$ge_winner <- factor(cont_winner2$ge_winner, levels = c("W", "N"))

levels(cont_winner2$ge_winner)
```

```{r, message=FALSE, warning=FALSE, echo=TRUE}
positions <- c("DEM","REP","DFL","UNK","OTH") #This will change the positon of the factors on the plot 

party_plot_bar <- ggplot(cont_winner2, aes(x = cand_party_affiliation, y = transaction_amt, fill = ge_winner)) +
  geom_bar(position = "dodge", stat = "identity") + 
  labs( x = "Party Affiliation", y = "Transaction Amount (millions)") + 
  scale_fill_discrete(name = "Election Results", labels = c("Win", "Lose")) +
  scale_x_discrete(limits = positions) +
  scale_y_continuous(limits = c(0, 2500000)
, breaks = seq(0, 2500000, 500000), expand = c(0,0), labels = c("0M","0.5M","1M","1.5M","2M","2.5M")) + 
  labs(title = "Does transaction amount affect win rate?", subtitle = "Exploring the relationship between contributions and election results") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_economist_white() + scale_color_economist()

party_plot_bar
```

The data graphic is a bar graph that illustrates total transaction amounts for 5 parties with the highest membership, split between wins and losses. 

The most popular 5 parties are Democratic party, Republican party, Minnesota Democratic-Farmer-Labor-Party, Unknown, and Others. The graphic concludes that there is not enough information to conclude that high financial contributions does not necessarily contribute to a win. Data wrangling concludes that our hypothesis was false, One can observe there is no relationship shown between election results and contributions Losses under the Republican party, for example, show higher total contribution of 1 million dollars versus total transaction amounts for Republican wins show a total contribution of about 800,000 thousand dollars. On the Other hand the Democratic party shows that the total transaction for candidates who won is 2.3 million in contributions versus transactions for Democratic losses of 1.3 million.


